Merow 2014 Appendix

### 2018-02-23 Merow2014_Appendix.Rmd
Appendix E - good & bad years; Appendix F - environmental covariate

>>>>

```{r App_E.1}
########
## simulate IPM matrices
########
  dff <- generateData()
  gr1 <- makeGrowthObj(dff, sizeNext ~ size + size2)
  sv1 <- makeSurvObj(dff, surv ~ size + size2)
  fv1 <- makeFecObj(dff, Transform="log")
  Pmat <- makeIPMPmatrix(nBigMatrix=100, minSize=-3, maxSize=14,
                         growObj=gr1, survObj=sv1, correction="constant")
  Fmat <- makeIPMFmatrix(nBigMatrix=100, minSize=-3, maxSize=14,
                         fecObj=fv1, correction="constant")
  IPM <- Pmat
  IPM@.Data <- Pmat+Fmat  # retains Pmat meshpoints, etc

  w.eigen <- Re(eigen(IPM)$vectors[,1])
  stable.dist <- w.eigen/sum(w.eigen)

########
## generate starting population
#########
  # size distribution
  size0 <- runif(1000, min(IPM@meshpoints), max(IPM@meshpoints))
  # IPM breakpoints = (mdpts - h/2):(max(mdpts) + h/2)
  breakpoints <- c(IPM@meshpoints - (IPM@meshpoints[2] - IPM@meshpoints[1])/2,
                   IPM@meshpoints[length(IPM@meshpoints)] +
                     (IPM@meshpoints[2] - IPM@meshpoints[1])/2)
  # bin size0 based on breakpoints
  new.dist.i <- findInterval(size0, breakpoints, all.inside=TRUE)
  loc.sizes <- table(new.dist.i)
  new.dist.n0 <- rep(0, length(IPM@meshpoints))
  new.dist.n0[as.numeric(names(loc.sizes))] <- loc.sizes

########
## project population forward
########
  tmax <- 10
  storage <- matrix(NA, nrow=dim(Pmat)[1], ncol=tmax)
  new.dist.n <- new.dist.n0
  for(k in 1:tmax) {
    new.dist.n <- IPM %*% new.dist.n
    storage[,k] <- new.dist.n
  }

########
## plots
########
  par(mfrow=c(2,2), mar=c(4,4,2,1))
  image.plot(Pmat@meshpoints, Pmat@meshpoints, t(Pmat), 
             main="P", xlab="Size (t)", ylab="Size (t+1)")
  contour(Pmat@meshpoints, Pmat@meshpoints, t(Pmat), add=TRUE)
  image.plot(Pmat@meshpoints, Pmat@meshpoints, t(Fmat), 
             main="F", xlab="Size (t)", ylab="")
  contour(Pmat@meshpoints, Pmat@meshpoints, t(Fmat), add=TRUE)
  hist(size0, col="grey", xlim=range(Pmat@meshpoints), freq=FALSE,
       xlab="Size (t=0)", ylab="", main="")
  for(k in 1:tmax) {
    lines(Pmat@meshpoints, storage[,k]/sum(storage[,k]), col=rgb(0,0,k/tmax,1))
  }
  lines(Pmat@meshpoints, stable.dist, col="red")
  image.plot(1:tmax, Pmat@meshpoints, t(log(storage)), 
             xlab="time", ylab="Size")
  abline(v=1.5+(1*c(0:tmax)))
  Re(eigen(IPM)$values[1])
```

```{r App_E.2}
dff <- generateData()

# good year
gr1 <- makeGrowthObj(dff, Formula=sizeNext ~ size, 
                     regType="constantVar", Family="gaussian")
sv1 <- makeSurvObj(dff, Formula=surv ~ size + size2)
fv1 <- makeFecObj(dff, Formula=list(fec~size), Transform="log")
fv1@fitFec[[1]]$coefficients <- c(-1, 0.35)

# bad year
sv1 <- coerceSurvObj(survObj=sv1, coeff=c(-1, 1, 2))
sv2 <- coerceSurvObj(survObj=sv1, coeff=c(-3, 0, -2))
gr2 <- coerceGrowthObj(gr1, coeff=c(-3, -3), sd=1)

# kernel
Pmat1 <- makeIPMPmatrix(nBigMatrix=50, minSize=-2, maxSize=12,
                        growObj=gr1, survObj=sv1)
Pmat2 <- makeIPMPmatrix(nBigMatrix=50, minSize=-2, maxSize=12,
                        growObj=gr2, survObj=sv2)
Fmat <- makeIPMFmatrix(nBigMatrix=50, minSize=-2, maxSize=12, fecObj=fv1)
IPM1 <- Pmat1 + Fmat
IPM2 <- Pmat2 + Fmat

# plots
image.plot(Pmat1@meshpoints, Pmat1@meshpoints, t(Pmat1), main="good year")
image.plot(Pmat1@meshpoints, Pmat1@meshpoints, t(Pmat2), main="bad year")

exp(stochGrowthRateSampleList(list(IPM1, IPM2), nRunIn=1000, tMax=5000))

nsim <- 100
tmax <- 100
IPM.ls <- list(IPM1, IPM2)
nt0 <- abs(Re(eigen(IPM1)$vector[,1])*10)
rt <- rtGood <- rtBad <- matrix(NA, nsim, tmax)
for(j in 1:nsim) {
  ntGood <- ntBad <- nt <- nt0
  for(k in 1:tmax) {
    sv.t <- coerceSurvObj(sv1, coeff=runif(3, -3, 3))
    gr.t <- coerceGrowthObj(gr1, coeff=runif(2, -3, 3), sd=1)
    Pmat.t <- makeIPMPmatrix(nBigMatrix=50, minSize=-2, maxSize=12,
                        growObj=gr.t, survObj=sv.t)
    IPM.t <- Pmat.t + Fmat
    nt <- IPM.t %*% nt
    rt[j,k] <- sum(nt)
    ntGood <- IPM1 %*% ntGood
    rtGood[j,k] <- sum(ntGood)
    ntBad <- IPM2 %*% ntBad
    rtBad[j,k] <- sum(ntBad)
  }
}

par(mfrow=c(1,1), bty="l")
matplot(t(rt), type="l", xlab="time", ylab="N", log="y", 
        ylim=c(range(c(rtGood, rtBad))), lty=1, col=rgb(0,0,0,0.5))
points(rtGood[1,], type="l", lwd=2, lty=2)
points(rtBad[1,], type="l", lwd=2, lty=2)
```

```{r App_E.3}
dff <- generateData()
go <- makeGrowthObj(dff, sizeNext ~ size + size2 + covariate)
so <- makeSurvObj(dff, surv ~ size + size2)
eo <- makeEnvObj(dff)
fo <- makeFecObj(dff, Formula=list(fec~size))

Pmat <- makeCompoundPmatrix(nEnvClass=2, envMatrix=eo, 
                            minSize=-3, maxSize=14, nBigMatrix=100,
                            growObj=go, survObj=so, 
                            correction="constant")
image.plot(1:nrow(Pmat), 1:ncol(Pmat), t(log(Pmat)), axes=F, 
           xlab="Size (t)", ylab="Size (t+1)")
axis(1, at=1:nrow(Pmat), lab=round(rep(Pmat@meshpoints, Pmat@nEnvClass)))
axis(2, at=1:nrow(Pmat), lab=round(rep(Pmat@meshpoints, Pmat@nEnvClass)))
abline(h=length(Pmat@meshpoints)*(1:Pmat@nEnvClass))
abline(v=length(Pmat@meshpoints)*(1:Pmat@nEnvClass))
```

```{r App_F}
########
## load data
########
  load("data/ActaeaRegDat.RData")
  d <- ActaeaRegDat

########
## regressions
########
  g.lm <- lm(y ~ x + K, data=d)
  s.lm <- glm(surv ~ x, family="binomial", data=d)
  fl.lm <- glm(flow ~ x, family="binomial", data=d)
  fr.lm <- glm(fruits ~ x, family="poisson", data=d)

  par(mfrow=c(2,2), mar=c(4,4,2,1))
  xx <- data.frame(x=seq(2, 14, length.out=100), K=25)
  plot(d$x, d$surv, col=rgb(0,0,0,0.1), xlab="Size (t)", ylab="Survival to t+1")
  lines(xx$x, predict(s.lm, xx, type="response"), col="red", lwd=2)
  plot(d$x, d$y, xlab="Size (t)", ylab="Size (t+1)")
  lines(xx$x, predict(g.lm, xx, type="response"), col="red", lwd=2)
  plot(d$x, d$flow, col=rgb(0,0,0,0.1), xlab="Size (t)", ylab="Pr(Flower)")
  lines(xx$x, predict(fl.lm, xx, type="response"), col="red", lwd=2)
  plot(d$x, d$fruits, xlab="Size (t)", ylab="Fruit number")
  lines(xx$x, predict(fr.lm, xx, type="response"), col="red", lwd=2)

  p <- rep(0, 15)
  p <- list("s.b0"=coef(s.lm)[1], "s.bx"=coef(s.lm)[2], 
         "g.b0"=coef(g.lm)[1], "g.bx"=coef(g.lm)[2], "g.bK"=coef(g.lm)[3],
         "fl.b0"=coef(fl.lm)[1], "fl.bx"=coef(fl.lm)[2],
         "fr.b0"=coef(fr.lm)[1], "fr.bx"=coef(fr.lm)[2],
         "est.x.mn"=3.0721075, "est.x.sd"=1.448899,
         "sd.xy"=1.155625, 
         "est.sdlg"=0.0091550, "surv.sdlg"=0.2383389, 
         "seed.pred"=0.195905)

########
## define functions
########
  # survival probability
  sx <- function(x, p) {
    u <- exp(p$s.b0 + p$s.bx*x)
    return(u/(1+u))
  }

  # growth
  gyx <- function(y, x, K, p) {
    dnorm(y, p$g.b0 + p$g.bx*x + p$g.bK*K, p$sd.xy)
  }

  # fecundity
  fx <- function(x, p) {
    # flowering probability
    u <- exp(p$fl.b0 + p$fl.bx*x)
    # fruits per flowering individual
    n.fr <- exp(p$fr.b0 + p$fr.bx*x)
    n.fr[n.fr<0] <- 0
    n.seed.per.fruit <- p$seed.pred*0.38901 + (1 - p$seed.pred)*9.61368
    return(u/(1+u)*n.fr*n.seed.per.fruit)
  }

  # seedling survival and size distribution
  py3 <- function(y, p) {
    return(p$surv.sdlg * dnorm(y, p$est.x.mn, p$est.x.sd)/
             (1-pnorm(minsize, p$est.x.mn, p$est.x.sd)))
  }

  # iterator function
  iter <- function(Nt) {
    Nt1 <- array(0, dim=c(2+n.size)) 
    Nt1[1] <- Nt1[1] + sum(FF %*% Nt[3:(n.size+2)])
    Nt1 <- Nt1 + as.vector(P %*% Nt)
    return(Nt1)
  }

########
## construct kernel
########
  # define breakpoints, meshpoints, stepsize
  n.size <- 100
  tol <- 1.e-8
  minsize <- 0.9234274; maxsize <- 15.27885
  L <- minsize; U <- maxsize; n <- n.size
  b <- L + c(0:n)*(U-L)/n
  y <- 0.5*(b[1:n]+b[2:(n+1)])
  h <- y[2]-y[1]

  # loop over forest succession (=approx(cells))
  spruce <- 0
  K.conc <- 0
  for(k in 1:120) {
    # spruce proportion & K concentration
    spruce <- 1/(1+exp(-0.09*k+5))
    K.conc <- exp(3.00932 - 0.53267*spruce)*1.722324
    # kernel components
      ## growth (including eviction correction)
      G <- array(0, dim=c(n.size, n.size))
      G <- h*outer(y, y, gyx, p=p, K=K.conc)
      for(i in 1:(n.size/2)) G[1,i] <- G[1,i] + 1 - sum(G[,i])
      for(i in (n.size/2+1):n.size) G[n.size,i] <- G[n.size,i] + 1 - sum(G[,i])
      ## survival
      S <- sx(y, p)
      FF <- fx(y, p)*h
      P <- array(0, dim=c(2+n.size, 2+n.size)) # [1,] = seed; [2,] = seedling
      P[2,1] <- p$est.sdlg
      P[3:(2+n.size),2] <- py3(y, p)
      for(i in 3:(2+n.size)) P[i,3:(2+n.size)] <- S*G[(i-2),]

    # population iterations
    Nt <- array(1, dim=c(2+n.size))
    Nt <- Nt/sum(Nt)
    qmax <- 1000; lam <- 1
    while(qmax > tol) {
      Nt1 <- iter(Nt)
      qmax <- sum(abs(Nt1 - lam*Nt))
      lam <- sum(Nt1)
      Nt <- Nt1/lam
    }
    stable.dist <- Nt/sum(Nt)
    stable.lam <- lam
    cat("Year:", k, "Proportion spruce:", round(spruce,2), " Lambda:", lam, "\n")
  }
```
